Molecular dynamics simulation of synchronization in driven particles

### 2020-12-30 manuscript/manuscript_ajp_2020.tex
found a sign error!

\begin{abstract}
  %What problem did you study and why is it important?
  Particles
  %confined to a narrow channel 
  driven
  across a washboard potential energy landscape
  exhibit a variety of synchronization effects
  in
  experimental systems 
  %interest to condensed matter physics 
  including 
  magnetically driven colloidal particles
  confined by light-fields.
  %  Numerical model
  We discuss a simple
  numerical model that
  %exhibits
  provides insight
  to synchronization behavior
  %and
  %atoms on surfaces.
  %voltage driven superconducting vortices
  %confined in a Josephson junction,
  %and ac and dc driven
  %charge and spin density waves.
  %We present the basics
  using
  molecular dynamics simulations
  for particles moving through
  a viscous liquid.
  %applied to single and multiple particle systems.
  %so we model the dynamics of each particle
  %with 
  %overdamped equations of motion.
  We demonstrate how to 
  visualize
  the system with
  numerical representations of the landscape
  and
  animations of the particle motion.
  Our results show a variety of systems that
  synchronize 
  %and detect synchronization using
  %measurements of
  particle velocity as a function of applied driving force,
  and characterize these behaviors
  with force-velocity curves and phase diagrams.
  We include sample code
  and exercises for students
  %that include
  with 
  opportunities
  to reproduce our results and propose
  new numerical experiments.
  %With only a few particles in two-dimensions,
  %the simulation runs quickly,
  %making this an appropriate model for undergraduates to explore.

  %Hew to this recipe (MMRC) witlessly
  %1. M—Give the motivation and context
  %2. M—Explain what you did in sufficient
  %detail so the audience knows if your work
  %is relevant and applicable to his or hers
  %3. R—Emphasize your key results
  %4. C—Tell the reader what you think the
  %results mean

\end{abstract}
% AJP requires an abstract for all regular article submissions.
% Abstracts are optional for submissions to the "Notes and Discussions" section.

@@

Experiments with colloid particles,
such as plastic beads suspended in
highly de-ionized water \cite{},
are useful model systems
to understand complex behaviors.
%silica spheres in organic solvents
%magnetic spheres?
%
Experiments of colloid particles confined
in optical traps
subject to external driving forces 
have
been used to examine the microscopic dynamics
of individual particles during mode-locking,
which cannot be easily imaged in electronics.
%experimental - colloids specifically
Colloidal particles trapped in light fields are %have proven
a useful medium for studying complex dynamical behaviors  %cite
due to their 
relatively large size of microns,
making them relatively simple 
to control and image.
Additionally colloids can mimic different
systems since 
the interaction forces between colloids can be modified
by tuning the chemistry of the suspending fluid or surface ligands
of the particles.
Because colloids are easy to control and study, 
they are used as experimental models
for systems relatively hard to access and visualize,
such as cold atoms or electron gases \cite{Grier2003}. %ELABORATE AND CITE
%Recent work with atomic systems such as Bose-Einstein condensates \cite{Zhang2020}
Mode-locking 
in colloids has been achieved in systems interacting via magnetic dipoles 
or electric forces 
in experiment \cite{Juniper2015, Lutz2004} and
simulation \cite{Herrera-Velarde2007, Herrera-Velarde2008}.

%-----------------------------------------------------------
%particle interaction types
%Particles which interact over long distances
%include colloids, magnetic beads, superconducting vortices, dusty plasmas, electron gases. [more%e detail and references]
%Particles which interact over short distances include
%bubble arrays/emulsions [more systems and references].----------------------
%The presence of a modulating surface
%can modify these patterns in a variety of ways,
%changing the onset of dynamical flows,
%and the overall flow patterns.
%  adjustable strength, geometry and length scale of substrate potential
%Baumgartl, Brunner, Bechinger, PRL 93, 168301 (2004).
%Baumgartl, Zvyagolskaya, Bechinger, PRL 99, 205503 (2007).

@@

%Given the prevalence of synchronization,
%it is the subject of many computational models.
Here we perform numerical studies 
on the synchronized dynamics
of confined particles driven over
a washboard shaped potential energy landscape.
We chose this model for its
relevance to condensed matter systems
and ease of simulation.
Numerical modeling of colloids can provide mechanistic insight
that can be difficult to achieve in experimental conditions
where Brownian motion and other sources of noise dominate.
%We examine
The period of the substrate
can be used to control the
intrinsic velocity of the dc driven particle,
and the applied ac drive can
create Shapiro steps in the force-velocity relationship.

@@

%-----------------------------------------------------------------
%transition from synchronized to chaotic
%describe types of physical systems - particle types
%%In this work,
%we perform 
%numerical simulations of confined, driven particles
%to model a variety of physical phenomena.
%-----------------------------------------------------------------------
%paper outline
%In the following paper
We describe
our molecular dynamics model for a single particle in Section~\ref{sec:MD}.
%In this section, we restrict our discussion to a single particle,
%and include additional information
%for multiple particles in Section~\ref{sec:sync}.
%including various  and confining environments.
We summarize
our results 
in Section~\ref{sec:results} 
including synchronized motion of a single confined particle
driven across a periodic landscape in 
Section~\ref{sec:one}.
In Section~\ref{sec:sync}
we 
extend the model to include multiple particles,
and show how confined 
interacting particles respond to the same applied driving force.
%including stationary propagation of high density kinks
%in Sec.~\ref{sec:kink}.
%We include code to simulate
%and visualize the landscape and dynamics in this section
%and supplementary material.
%we
%demonstrate how uniform environments and applied forces
%and create synchronized flow patterns.
%We present these results using standard tools of non-linear oscillators
%such phase diagrams of velocity versus position.
%PHASE DIAGRAMS SHOWING COUPLING, OR QUASI-PERIODIC COUPLING
%INDICATE COOPERATIVE MOTION IN THE Q1D SUBSTRATE
%In Section~\ref{sec:quasiperiod},
%we show how an aperiodic landscape modifies the particle dynamics.
%Finally we explore the transition to chaotic dynamics in 
%Section~\ref{sec:chaos}
%and conclude 
%in Section~\ref{sec:conclusion}. 
We conclude with
exercises for interested students 
in Section~\ref{sec:problems}.

@@

\begin{center}
\begin{figure}[h!]
\centering
\includegraphics[width=\columnwidth]{landscape.pdf}
\caption{
  Schematic of the simulation
  of a single particle
  driven across a washboard potential
  energy landscape.
  (a) View of the $x-y$ plane. %of the simulation
  The time-dependent applied driving force $\vec{F}^D$
  is parallel to the y-axis.
  The force due to the landscape %$\vec{F}^l$
  is calculated due to the gradient of the potential
  energy 
  $\vec{F}^l = -\nabla U(\vec{r})$.
  (b) The potential energy function
  along the y-axis $U_l(y)$
  with period $\lambda = L/N_p$
  where the number of periods is $N_p = 3$.
  }
\label{fig:landscape0}
\end{figure}
\end{center}

We use a classical two-dimensional model for 
studying the dynamics of $N$ interacting particles,
using the net force on each particle to calculate
its trajectory.
Particles are confined in a two-dimensional (2D) 
simulation of area $A = L \times L$ where $L=36.5 a_0$
where $a_0$ is a dimensionless unit of length.
An individual particle $i$ has
position $\vec{r}_i = x_i \hat{x} + y_i \hat{y}$
and velocity $\vec{v}_i = d\vec{r}_i/dt$.
The edges of the system are treated with
periodic boundary conditions
such that a particle leaving the edge of the system is mapped
back to a position within the simulation 
by the transformation $x_i+L \rightarrow x_i$ and $y_i+L \rightarrow y_i$.
%and particles interact across PBC.
We show a schematic of the system in Fig.~\ref{fig:0}

We confine the particles using a position dependent 
potential energy function, called a landscape or substrate.
The confining force %arises from a potential function 
 %creating a landscape of potential minima and maxima
%that modify the force on a particle as a function of position
is calculated as 
 $\vec{F}^{l}_i(\vec{r}_i) = -\nabla U_l(\vec{r}_i)$.
 %The landscape potential energy $U(\vec{r})$ is static 
 %with fixed minima and maxima
 %that are periodic or quasi-periodic,
 %as described in Sec.~\ref{sec:results}.
% 
The periodic landscape is modulated in the y-direction as
%% %that used in Sec.~\ref{sec:one}
% where 
 \begin{equation}
   U(y) = U_0 \cos{(N_p \pi y / L)}
     \label{eq:ysubstrate}
\end{equation}
 where $N_p$ are the number of periods,
 % $L$ is the period length,
 and $U_0$ is an adjustable parameter
 to set the depth of the troughs.
 We define the substrate strength to be
 the maximum force the substrate can exert on a particle 
 $A_p = 2\pi U_0/L$.
 %

 Particles are subject an external time-dependent driving force
$\vec{F}_{D}(t)$
applied parallel to the y-direction.
We model this force as
\begin{equation}
  \vec{F}^{D}(t) = [F^{DC} + F^{AC} \sin(\omega t)] \hat{y},
    \label{eq:drive}
\end{equation}
with modifiable parameters including
a constant component $F^{DC}$,
and a time dependent component with amplitude $F^{AC}$
and frequency $\omega = 2 \pi f$.
%The frequency is scaled in time units,
%where [finish].
When $F^{DC}$ is non-zero,  
%In several simulations
we slowly increase 
$F^{DC}$ from zero
to avoid transient effects.

%because 
We model the 
particle dynamics 
with an overdamped
equation of motion.
%integrated with the Verlet method.
%The suspending fluid is highly viscous
%and exerts a damping force on the particles equivalent
%to a linear drag force.
%Overdamped dynamics is a common assumption in models of colloidal particles.
%This is an assumption that
The term `overdamped or critically damped indicates
that an oscillator cannot undergo periodic
undulations due to some external force. %such as
%suspension in a viscous fluid \cite{}.
In our colloidal model we assume
the damping comes from the viscosity of the fluid,
where a sphere of radius $0.5 \mu m$ has a Reynolds number
of $R \approx 10^{-6}$, indicating that linear drag dominates its motion.
See Section~\ref{sec:problems} Exercise~\ref{ex:reynolds} for more information
about the Reynold's number.
We model this
nonconservative force,
with the linear friction $\vec{F}_{drag} = -\eta \vec{v}_i$
sufficient so that
the acceleration of the particle is zero.
Such a model is appropriate when the particles
are small and the viscosity of the suspending medium is high.
% [finish... discussion of Reynolds number]
This model should be familiar to readers
modeling Millikan's oil drops
in a standard classical mechanics text \cite{Taylor2005}.
%This model is encountered in standard texts,
%for instance Ch. 2 of Taylor's Classical Mechanics.
%more precise: 6\pi \eta a \vec{v}

Newton's second law for an individual particle
$m \vec{a}_i = \sum \vec{F}$ is simplified
by the assumption $\vec{a}_i$ is zero in overdamped model.
%Newton's second law can be rearranged %for the velocity
%to
The equation of motion for an isolated particle 
\begin{equation}
  \eta \vec{v}_i = \vec{F}^l_{i} + \vec{F}^{D}(t).
    \label{eq:motion}
\end{equation}
where the viscosity coefficient $\eta = 1$.
In Section~\ref{sec:problems} Exercise~\ref{ex:n2l}
we invite the reader to confirm this result.
The equation of motion provides a direct calculation of the velocity
of an individual particle from its location $\vec{r}_i$ %in the simulation
with respect to other particles $j$.

The molecular dynamics simulation is controlled by a $for()$ loop
 which runs from an initial to maximum integer number of time steps $t_i$.
Each integer time step interval $t_{i+1}-t_i$
represents a simulation time interval of $\Delta t$ 
simulation units.  %as described in Table 1
At each time step
we evaluate the net force on each particle as a function of its position
$\vec{r}_i(t)$
and then integrate
the equation of motion to move particles
to an updated position
$\vec{r}_i(t+\Delta t)$.
%
Since the acceleration as zero,
[more background here - do the students necessarily know the Verlet method?  no]
the Verlet method simplifies to 
the Euler method,
which is used to calculate the position at subsequent time steps.
%Most undergraduate students learn the Euler method in Computational Physics and Differential Equations, but don't encounter the Verlet method.
\begin{equation}
  \vec{r}_i(t+\Delta t) = \vec{v}_i(t) \Delta t + \vec{r}_i(t).
    \label{eq:euler}
\end{equation}

The units of the simulated variables are summarized in Table ~\ref{tab:1}.

\begin{table}[h!]
\centering
\caption{Simulation units.}
\begin{ruledtabular}
\begin{tabular}{c c p{5cm}}
Quantity & Unit \\
\hline
length &  $a_0$ \\
energy & $E_0$ \\
%electric potential & $V(r_{ij}) = E_0/r_{ij}$ \\
%energy & $E_0 = q^2{Z^*}^2/4\pi \epsilon \epsilon_0 a_0$ \\
%dimensionless interaction strength & $q$ \\
%effective colloidal charge & $Z^*$ \\
%solvent dielectric constant & $\epsilon \epsilon_0$\\
force & $E_0 / a_0$\\
viscosity/damping constant & $\eta$ \\
time &  $\eta / E_0$ \\
velocity &  $ (E_0 / \eta) a_0$ \\
\end{tabular}
\end{ruledtabular}
\label{tab:1}
\end{table}

\section{Results}
\label{sec:results}
 We demonstrate how a single particle (Sec.~\ref{sec:one})
 and many particles
 move (Sec.~\ref{sec:sync})
 in response to this applied force in a variety of environments.
 In Sec.~\ref{sec:kink} we set $F^{DC}$
 to zero and track the motion
 of a high density area of a particle chain
 (i.e. kink dynamics).
 %The driving force does add energy into the system, and some of it is lost.
%
\subsection{Single particle system}
% In subsection headings, only the first word is capitalized.
\label{sec:one}

We drive a single particle across a periodic landscape 
along the y-direction 
\begin{equation}
  \label{eq:landscape_potential}
  U_l(y) = U_{0y} \cos{(N_p \pi y / L)} 
\end{equation}
with $N_p=20$ troughs in the landscape
and landscape period $\lambda = L/N_p = 1.825 a_0$.
%({\it redundant, just refer back to eq. in Model?})
%and $V_{0y}=0.2$.
This is illustrated in Fig.~\ref{fig:1}(a) where
the red (blue) regions show local maxima (minima).
The code for generating
a two dimensional colored plot
of the potential landscape
is calculated by evaluating
the analytic function in Eq.~\ref{eq:landscape}
for a grid of values $(x_n,y_n)$.
%is included in the supplementary material,
The numerical implementation of the landscape 
is calculated with
\begin{equation}
  \label{eq:force}
  F_y(y) = A_{p} \sin{(N_p \pi y / L)} 
\end{equation}
where $A_{p} = 0.1$.

Because we are performing molecular dynamics
for a single particle on a smooth potential energy landscape,
we use a large simulation time step of 
$\Delta t = 0.01$.
In Sec.~\ref{sec:sync},
we decrease the time step
to treat multiple particle interactions.

We drive the particle with
$F^{AC}=0.2$, $F^{DC}=0.01$, and $f=0.01$ cycles per time unit.
When the superposition of $F^{AC}$ and $F^{DC}$ is large enough
to overcome the barrier height of the landscape troughs,
the particle hops between troughs in the energy landscape.
In Fig.~\ref{fig:0} we
show the relationship between the applied force and particle
position.
In Fig.~\ref{fig:0}(a)
we plot the applied driving force
as a function of time
and in Fig.~\ref{fig:0}(b) 
we show the position of the particle
as a function of time.
The initial position is near $y=0$
and the initial driving force is minimum
$F_D(t=0) = F^{DC}$.
Over the time period $T/2 = 50$ %time units
the particle moves
in the positive y-direction
while $F_{D}(t) > 0 $
through three
substrate troughs, reaching a maximum of $y/\lambda = 3$.
When $F_{D}(t) < 0 $
the particle moves in the negative y-direction
reaching a minimum position
of $y/\lambda = 1$
over the period $T = 1/f = 100$.
The average velocity $\bar{v}_y$
is the average displacement $\Delta y = y(t_0+T) - y(t_0)$
over the period of the driving force.
In Fig. ~\ref{fig:0}
the average displacement during one time period is a single wavelength
of the substrate $y(t_0+T) - y(t_0) = \lambda$.
Thus the average velocity 
is $\bar{v}_y = \lambda f$. %, where $T = 1/f$.

@@

>>>>

\begin{abstract}
  %What problem did you study and why is it important?
  Particles
  %confined to a narrow channel 
  driven
  across a washboard potential energy landscape
  exhibit a variety of synchronization effects.
  %in
  %  Numerical model
  We discuss a simple
  numerical model that
  %exhibits
  provides insight
  to synchronization behavior
  %and
  %atoms on surfaces.
  %voltage driven superconducting vortices
  %confined in a Josephson junction,
  %and ac and dc driven
  %charge and spin density waves.
  %We present the basics
  using
  molecular dynamics simulations
  for particles moving through
  a viscous liquid.
  This model reproduces behaviors observed
  in experimental systems 
  of interest to condensed matter physics 
  including 
  magnetically driven colloidal particles
  confined by light-fields.
  %
  %applied to single and multiple particle systems.
  %so we model the dynamics of each particle
  %with 
  %overdamped equations of motion.
  We demonstrate how to 
  visualize
  the system with
  numerical representations of the landscape
  and
  animations of the particle motion.
  Our results show a variety of synchronization effects 
  in single and multi-particle systems, 
  %and detect synchronization using
  %measurements of
  which we characterize with
  plots of 
  particle velocity as a function of applied driving force,
  and phase diagrams of position versus velocity.
  %We include sample code
  %and exercises for students
  %with 
  %opportunities
  %to reproduce our results and propose
  %new numerical experiments.
  %With only a few particles in two-dimensions,
  %the simulation runs quickly,
  %making this an appropriate model for undergraduates to explore.  
\end{abstract}
% AJP requires an abstract for all regular article submissions.
% Abstracts are optional for submissions to the "Notes and Discussions" section.

@@

Complex dynamics can be studied with
colloid particles with experiments and simulation.
In experiment, colloids are typically spheres of 
plastic suspended in
highly de-ionized water or silica in suspended in organic solvent.
%experimental - colloids specifically
The spheres are a few microns in size
and move at rates of microns per second,
so a system can be imaged
with a traditional optical camera
due to the large size and slow speed of the particles.
Individual colloids exhibit 
Brownian motion,
where particles change direction in a seemly random fashion
due to unseen 
collisions 
with particles comprising the suspending fluid \cite{}.
Brownian motion is a temperature dependent
phenomena affecting quantum mechanically small systems. \cite{}
[elaborate].

Due to the relative ease of
manipulation and imagining, 
colloids are used as experimental models
for systems relatively hard to access and visualize. 
%SAME SENTENCE TWICE
Collections of colloids can
be used to study the properties of solids and liquids
%comprised of smaller particles
%such as atoms \cite{} or electrons \cite{}
or more exotic systems 
such as cold atoms or electron gases \cite{Grier2003}.  
The interaction forces between colloids can be controlled 
by altering the chemistry of the suspending fluid
or surface ligands of the particles.
Thus colloids can 
be modified to mimic different
systems such as hard spheres interacting via contact forces \cite{},
dipole interactions \cite{},
and long range electrostatic interactions \cite{}.

Light is a
tool for manipulating the colloidal environment.
A single laser beam can be used to create an optical trap,
which uses 
photon radiation pressure to
control colloid motions.
A circular optical trap is designed so the center is
a local energy minimum where 
a colloid in the beam center is subject to a
uniform distribution of photon bombardment. %, making 
Colloids at a trap edge
experience a net force
due to the uneven 
distribution of photon collisions on its surface.
Depending on the shape of the trap and location of the particle in the trap,
the radiation pressure either pushes colloids toward the local minima
or ejects it from the trap.
A single optically trapped colloid executing Brownian motion
is a useful probe of microscopic forces,
and has been studied in detail in Ref.~\cite{Volpe2013}.

Complex environments can be
engineered with light by %, typically made 
using diffraction gratings to create a patterned
potential energy landscape 
with one or more laser beams \cite{}.
%

Experiments of colloid particles confined
in optical traps arrays or patterns 
subject to external driving forces 
have
been used to examine the microscopic dynamics
of individual particles during mode-locking \cite{Juniper2015}.
This produced step-by-step diagrams
of particle dynamics that cannot be 
imaged in electronic systems such as Josephson junctions.
%Recent work with atomic systems such as Bose-Einstein condensates \cite{Zhang2020}
Thus the study of
mode-locking is a useful problem
of complex quantum mechanical systems.
Mode-locked 
colloid dynamics have been achieved in 
experiments \cite{Juniper2015, Lutz2004} and
simulations \cite{Herrera-Velarde2007, Herrera-Velarde2008} 
for a variety of interaction types such as 
magnetic dipoles and electric forces.
%-----------------------------------------------------------
%particle interaction types
%Particles which interact over long distances
%include colloids, magnetic beads, superconducting vortices, dusty plasmas, electron gases. [more%e detail and references]
%Particles which interact over short distances include
%bubble arrays/emulsions [more systems and references].----------------------
%The presence of a modulating surface
%can modify these patterns in a variety of ways,
%changing the onset of dynamical flows,
%and the overall flow patterns.
%  adjustable strength, geometry and length scale of substrate potential
%Baumgartl, Brunner, Bechinger, PRL 93, 168301 (2004).
%Baumgartl, Zvyagolskaya, Bechinger, PRL 99, 205503 (2007).

@@

Numerical modeling of colloids can provide mechanistic insight
that can be difficult to achieve in experimental conditions
where Brownian motion and other sources of noise dominate.

%Given the prevalence of synchronization,
%it is the subject of many computational models.
Here we perform numerical studies 
on the synchronized dynamics
of confined particles driven over
a washboard shaped potential energy landscape.
We chose this model for its
relevance to condensed matter systems
and ease of simulation.
%We examine
The period of the substrate
can be used to control the
intrinsic velocity of the dc driven particle,
and the applied ac drive can
create Shapiro steps in the force-velocity relationship.

@@

%-----------------------------------------------------------------
%transition from synchronized to chaotic
%describe types of physical systems - particle types
%%In this work,
%we perform 
%numerical simulations of confined, driven particles
%to model a variety of physical phenomena.
%-----------------------------------------------------------------------
%paper outline
%In the following paper
We describe
our molecular dynamics model for a single particle in Section~\ref{sec:MD}.
%In this section, we restrict our discussion to a single particle,
%and include additional information
%for multiple particles in Section~\ref{sec:sync}.
%including various  and confining environments.
We summarize
our results 
%in Section~\ref{sec:results} 
including synchronized motion of a single confined particle
driven across a periodic landscape in 
Section~\ref{sec:one}.
In Section~\ref{sec:sync}
we 
extend the model to include multiple particles,
and show how confined 
interacting particles respond to the same applied driving force.
%including stationary propagation of high density kinks
%in Sec.~\ref{sec:kink}.
%We include code to simulate
%and visualize the landscape and dynamics in this section
%and supplementary material.
%we
%demonstrate how uniform environments and applied forces
%and create synchronized flow patterns.
%We present these results using standard tools of non-linear oscillators
%such phase diagrams of velocity versus position.
%PHASE DIAGRAMS SHOWING COUPLING, OR QUASI-PERIODIC COUPLING
%INDICATE COOPERATIVE MOTION IN THE Q1D SUBSTRATE
%In Section~\ref{sec:quasiperiod},
%we show how an aperiodic landscape modifies the particle dynamics.
%Finally we explore the transition to chaotic dynamics in 
%Section~\ref{sec:chaos}
%and conclude 
%in Section~\ref{sec:conclusion}. 
We conclude with
exercises for interested students 
in Section~\ref{sec:problems}.

@@

We use a classical model for 
studying the dynamics of $N$ interacting particles,
using the net force on each particle to calculate
its trajectory.
Particles are confined in a two-dimensional (2D) 
simulation of area $A = L \times L$ where $L=36.5 a_0$
where $a_0$ is a dimensionless unit of length.
An individual particle $i$ has
position $\vec{r}_i = x_i \hat{x} + y_i \hat{y}$
and velocity $\vec{v}_i = d\vec{r}_i/dt$.
The edges of the system are treated with
periodic boundary conditions
such that a particle leaving the edge of the system is mapped
back to a position within the simulation 
by the transformation $x_i+L \rightarrow x_i$ and $y_i+L \rightarrow y_i$.
%and particles interact across PBC.
We show a schematic of the system in Fig.~\ref{fig:landscape0}(a).

We confine the particles using a position dependent 
potential energy function, called a landscape or substrate.
The confining force %arises from a potential function 
 %creating a landscape of potential minima and maxima
%that modify the force on a particle as a function of position
is calculated as 
 $\vec{F}^{l}_i(\vec{r}_i) = -\nabla U_l(\vec{r}_i)$.
 %The landscape potential energy $U(\vec{r})$ is static 
 %with fixed minima and maxima
 %that are periodic or quasi-periodic,
 %as described in Sec.~\ref{sec:results}.
% 
The landscape is modulated in the y-direction
with the periodic function 
%% %that used in Sec.~\ref{sec:one}
% where 
 \begin{equation}
   U(y) = U_0 \cos{(N_p \pi y / L)}
     \label{eq:ysubstrate}
\end{equation}
 where $N_p$ are the number of periods,
 % $L$ is the period length,
 and $U_0$ is an adjustable parameter
 to set the depth of the minima
 with simulation units of energy $E_0 = 1$. 
 %We define the substrate strength to be
 %the maximum force the substrate can exert on a particle 
 %$A_p = 2\pi U_0/L$.
 %
 We plot this function in 
 Fig.~\ref{fig:landscape0}
 we use $N_p = 3$ to illustrate the features of
 Eq.~\ref{eq:ysubstrate}.
 In Fig.~\ref{fig:landscape0}(a) we show 
 the $x-y$ plane with a contour plot
 of the the 2D potential energy landscape,
 where 
 the maxima are colored red and the minima colored blue.
 In Fig.~\ref{fig:landscape0}(b) we plot the function
 $U(y)$ to illustrate how the magnitude
 $\vec{F}^{l}_i$ is calculated from particle position in the $y-$direction.

\begin{center}
\begin{figure}[h!]
\centering
\includegraphics[width=\columnwidth]{landscape.pdf}
\caption{
  Schematic of the simulation
  of a single particle
  driven across a washboard potential
  energy landscape.
  The period of the landscape, $\lambda = L/N_p$ 
  where $N_p = 3$, is marked in red.
  The force due to the landscape %$\vec{F}^l$
  is calculated from the gradient of the potential
  energy 
  $\vec{F}^l = -\nabla U(\vec{r})$.
  (a) View of the $x-y$ plane. %of the simulation
  The time-dependent applied driving force $\vec{F}^D$
  is parallel to the y-axis.
  The landscape is shown with a contour plot,
  with maxima in the potential energy marked in red
  and minima marked in blue.
  A particle is shown in the region between minima and maxima
  subject to competing forces due to the landscape and applied driving force.
  (b) The potential energy function
  along the y-axis $U_l(y)$,
  The particle in (a) is shown at the same $y-$position,
  where the magnitude of force $\vec{F}^l$ is calculated as the slope of this function,
  directed along the $y-$direction in panel (a).
  }
\label{fig:landscape0}
\end{figure}
\end{center}

 Particles are subject an external time-dependent driving force
$\vec{F}_{D}(t)$
applied parallel to the y-direction.
We model this force as
\begin{equation}
  \vec{F}^{D}(t) = [F^{DC} + F^{AC} \sin(\omega t)] \hat{y},
    \label{eq:drive}
\end{equation}
with modifiable parameters including
a constant component $F^{DC}$,
and a time dependent component with amplitude $F^{AC}$
and frequency $\omega = 2 \pi f$.
We illustrate this force in Fig.~\ref{fig:landscape0}(a)
in the positive $y-$direction.
%but it may 
%be zero or negative depending the the parameters and time.
When $F^{DC}$ is non-zero,  
%In several simulations
we slowly increase 
$F^{DC}$ from zero
to avoid transient effects.

%because 
We model the 
particle dynamics 
with an overdamped
equation of motion.
%integrated with the Verlet method.
%The suspending fluid is highly viscous
%and exerts a damping force on the particles equivalent
%to a linear drag force.
%Overdamped dynamics is a common assumption in models of colloidal particles.
%This is an assumption that
The term overdamped or critically damped indicates
that an oscillator cannot undergo periodic
undulations due to some external force. %such as
%suspension in a viscous fluid \cite{}.
In our colloidal model we assume
the damping comes from the viscosity of the fluid,
where a sphere of radius $0.5 \mu m$ has a Reynolds number
of $R \approx 10^{-6}$, indicating that linear drag dominates
the particle motion.
See Section~\ref{sec:problems} Exercise~\ref{ex:reynolds} for more information
about the Reynold's number.
In the overdamped model the
linear friction produced by this 
nonconservative force,
$\vec{F}_{drag} = -\eta \vec{v}_i$ is 
sufficient so that
the acceleration of the particle is zero.
Such a model is appropriate when the particles
are small and the viscosity of the suspending medium is high.

Newton's second law for an individual particle
$m \vec{a}_i = \sum \vec{F}$ is simplified
by the assumption $\vec{a}_i$ is zero in overdamped model.
%Newton's second law can be rearranged %for the velocity
%to
The equation of motion for an isolated particle is
%solved for velocity 
\begin{equation}
  \eta \vec{v}_i = \vec{F}^l_{i} + \vec{F}^{D}(t).
    \label{eq:motion}
\end{equation}
with the viscosity coefficient $\eta = 1$,
which we describe in more detail in Exercise~\ref{ex:reynolds}.
In Section~\ref{sec:problems} Exercise~\ref{ex:n2l}
we invite the reader to confirm this result.

The equation of motion provides a direct calculation of the velocity
of an individual particle from its location $\vec{r}_i$ %in the simulation
and the simulation time.
The molecular dynamics simulation is controlled by a $for()$ loop
 which runs from an initial to maximum integer number of time steps $t_i$.
Each integer time step interval $t_{i+1}-t_i$
represents a simulation time interval of $\Delta t$ 
simulation units.  %as described in Table 1
At each time step
we evaluate the net force on each particle as a function of its position
$\vec{r}_i(t)$
and then integrate
the equation of motion to move particles
to an updated position
$\vec{r}_i(t+\Delta t)$.
%
Since the acceleration as zero,
the integration of the equation of motion
is performed via 
the Euler method 
\begin{equation}
  \vec{r}_i(t+\Delta t) = \vec{v}_i(t) \Delta t + \vec{r}_i(t)
    \label{eq:euler}
\end{equation}
for a small time step $\Delta t$.
In Section~\ref{sec:problems} Exercise~\ref{ex:euler}
we elaborate on
the numerical methods for 
solving differential equations,
and demonstrate how the Verlet method
simplifies to the Euler method when $\vec{a}=0$.
%
The units of the simulated variables are summarized in Table ~\ref{tab:1}.

\begin{table}[h!]
\centering
\caption{Simulation units.}
\begin{ruledtabular}
\begin{tabular}{c c p{5cm}}
Quantity & Unit \\
\hline
length &  $a_0 = 1$ \\
energy & $E_0 = 1$ \\
%electric potential & $V(r_{ij}) = E_0/r_{ij}$ \\
%energy & $E_0 = q^2{Z^*}^2/4\pi \epsilon \epsilon_0 a_0$ \\
%dimensionless interaction strength & $q$ \\
%effective colloidal charge & $Z^*$ \\
%solvent dielectric constant & $\epsilon \epsilon_0$\\
force & $E_0 / a_0$\\
viscosity coefficient & $\eta = 1$ \\
velocity &  $ E_0 / (\eta a_0) $ \\
time &  $(\eta / E_0) {a_0}^2$ \\
\end{tabular}
\end{ruledtabular}
\label{tab:1}
\end{table}

%\section{Results}
%\label{sec:results}
% We demonstrate how a single particle (Sec.~\ref{sec:one})
% and many particles
% move (Sec.~\ref{sec:sync})
% in response to this applied force in a variety of environments.
% In Sec.~\ref{sec:kink} we set $F^{DC}$
% to zero and track the motion
% of a high density area of a particle chain
% (i.e. kink dynamics).
 %The driving force does add energy into the system, and some of it is lost.
%
\section{Mode-locking of a single particle}
% In subsection headings, only the first word is capitalized.
\label{sec:one}

We drive a single particle across a periodic landscape 
along the y-direction described in Eq.~\ref{eq:ysubstrate}
%\begin{equation}
%  \label{eq:landscape_potential}
%  U_l(y) = U_{0y} \cos{(N_p \pi y / L)} 
%\end{equation}
with $N_p=20$ %troughs in the landscape
and landscape period $\lambda = 1.825 a_0$.
%({\it redundant, just refer back to eq. in Model?})
%and $V_{0y}=0.2$.
%This is illustrated in Fig.~\ref{fig:1}(a) where
%the red (blue) regions show local maxima (minima).
%
The numerical implementation of the landscape 
is calculated with
\begin{equation}
  \label{eq:force}
  F^l_y(y) = -A_{p} \sin{(N_p \pi y / L)} 
\end{equation}
where the force is scaled with parameter $A_{p} = 0.1$
in simulation units for force. % $E_0/a_0$.

%%HMMM... HAVE YOU INTRODUCED A PHASE SHIFT?

In Fig.~\ref{fig:0} we
show the relationship between the applied force and particle
position.
In Fig.~\ref{fig:0}(a)
we plot the applied driving force defined in Eq.~\ref{eq:drive}
as a function of time.
The time independent constants are 
$F^{AC}=0.2$, $F^{DC}=0.01$, and $f=0.01$ cycles per time unit.
The time period of the driving force is
$T = 1/f = 100$,
with the initial driving force at is minimum value 
$F^D(0) = F^{DC}$.

In Fig.~\ref{fig:0}(b) 
we show the $y-$position of the particle
as a function of time,
where we 
normalize $y$ by the spatial period of the washboard as
$y/\lambda$.
The initial particle position is $y=0$,
where the landscape force is minimum $F^l = 0$.
Subsequent substrate 
%CHECK!!! I THINK THIS IS WRONG!!!
%BY NOT INCLUDING THE SIGN YOU'VE DEFINED IT BACKWARDS!!!!!
%AND LOCATED THE INITIAL POSITION OF THE PARTICLE
%AT A LOCAL MAXIMA!!!!!
%The substrate maxima are located at integer multiples of $y/\lambda = 0, 1, 2 $ 
%and
minima are located at integer multiples $y/\lambda = 1/2, 3/2, 5/2, $ {\it et cetera}.
%This is apparent in the spacing of $y/\lambda$ in time,
%where the visual gaps between positions in Fig.~\ref{fig:0}(b) 
%at half integer multiples $y/\lambda = 1/2, 3/2, 5/2$ indicate regions the particle was moving slowly.
%measure/plot velocity?

When $F^D(t) > A_p$, a particle can 
overcome the barrier height of the landscape troughs,
and 
the particle hops between minima in the energy landscape.
In Fig.~\ref{fig:0} we demonstrate this relationship,
where over half of a driving period ($T = 50$) %time units
while $F_{D}(t) > 0 $ 
the particle moves
in the positive y-direction
through three
substrate troughs, reaching a maximum of $y/\lambda = 3$.
When $F_{D}(t) < 0 $
the particle moves in the negative y-direction
reaching a minimum position
of $y/\lambda = 1$
over the total period $T = 100$.
The average velocity $\bar{v}_y$
is the average displacement $\Delta y = y(t_0+T) - y(t_0)$
over the period of the driving force.
In Fig. ~\ref{fig:0}
the average displacement during one time period is a single wavelength
of the substrate $y(t_0+T) - y(t_0) = \lambda$.
Thus the average velocity 
is $\bar{v}_y = \lambda f$. %, where $T = 1/f$.

@@

\item
  Generally when solving differential equations,
  the Euler method is effective for solving linear equations,
  such as the one-dimensional 
  $f(y) = df/dy$ 
  but not higher order methods.

  The Verlet method simplifies to

  %FOLLOW UP WITH THIS!
  Because we are performing molecular dynamics
  for a single particle on a smooth potential energy landscape,
  we use a large simulation time step of 
  $\Delta t = 0.01$.
  In Sec.~\ref{sec:sync},
  we decrease the time step
  to treat multiple particle interactions.

  \label{ex:euler}

\item Drawing the contour

  The code for generating
a two dimensional colored plot
of the potential landscape
is calculated by evaluating
the analytic function in Eq.~\ref{eq:landscape}
for a grid of values $(x_n,y_n)$.
%is included in the supplementary material,
